A Note on Convergence Diagnostics for Multiple Imputation using Chained Equations (MICE)

### 2019-12-11 Simulation/Writeup/NoteConvergence.tex
Checked references

The convergence guidelines introduced in this paper are developed to be integrated into the \pkg{mice} environment \citep{mice} in \proglang{R} \citep{R}. This note therefore follows notation and conventions of \cite{mice}. Deviations from the `original' notation by \cite{rubin87} are described in \citep[\S~XYZ]{buur18}.

@@

\textbf{$\widehat{R}$}. The $\widehat{R}$ values for chain means and chain variances show equivalent trends, neither is there a clear difference between the four variables. Roughly speaking, we see that $\widehat{R}$ is higher for simulation conditions with a lower number of iterations, and approaches one for conditions with more iterations. Figure \ref{fig:conv} shows an initial dip, a small bump upwards, and then a steep decline. This decline \textbf{houdt aan} up-to iteration twenty, after which a more gradual decrease can be observed up-to iteration forty. The $\widehat{R}$ values are more or less stable after that. The conventional acceptable threshold for convergence, maximum $\widehat{R} < 1.2$, is reached in the simulation condition with three iterations, and conditions with six iterations or more. The widely used threshold $\widehat{R} < 1.1$ is met in conditions with ten or more iterations. The most recent recommended threshold is not reached within the 100 iterations considered in this simulation study. $\widehat{R} > 1.01$ for all simulation conditions (see online appendix for the full table of results). \textbf{Check is computing r ht is still correct!}

\textbf{Auto-correlation.} All observed auto-correlations (aggregated over the 1000 repetitions) are negative or close to zero. There are some differences between variables, and the ACs of chain variances seem lower than the ACs of chain means. The general trend in figure \ref{fig:conv} is that ACs decrease steeply as the number of iterations increase.  \textbf{The lowest AC observed is %-0.735 
at four iterations. Simulation conditions with a higher number of iterations have show a slow increase towards zero. Auto-correlations do not plateau completely within the range of simulation conditions considered in this study. It seems that ACs might become positive if more iterations would be allowed. The average absolute maximum ACs are negative up to 77 iterations for chain means, and for chain variances all iterations considered here have negative ACs.}

>>>>

The convergence guidelines introduced in this paper are developed to be integrated into the \pkg{mice} environment \citep{mice} in \proglang{R} \citep{R}. This note, therefore, follows notation and conventions of \cite{mice}. Deviations from the `original' notation by \cite{rubin87} are described in \citep[\S~2.2.3]{buur18}.

@@

\textbf{$\widehat{R}$}. The $\widehat{R}$ values for chain means and chain variances show equivalent trends, neither is there a clear difference between the four variables. Roughly speaking, we see that $\widehat{R}$ is higher for simulation conditions with a lower number of iterations, and approaches one for conditions with more iterations. Figure \ref{fig:conv} shows an initial dip, a small bump upwards, and then a steep decline. This decline \textbf{houdt aan} up-to iteration twenty, after which a more gradual decrease can be observed up-to iteration forty. The $\widehat{R}$ values are more or less stable after that. The \textbf{conventionally acceptable threshold for convergence, maximum $\widehat{R} < 1.2$,[CHECK IN INTRO]} is reached in the simulation condition with three iterations, and conditions with six iterations or more. The widely used threshold $\widehat{R} < 1.1$ is met in conditions with ten or more iterations. The most recent recommended threshold is not reached within the 100 iterations considered in this simulation study. $\widehat{R} > 1.01$ for all simulation conditions (see online appendix for the full table of results). \textbf{Check is computing r ht is still correct! Also say something about Rh below 1.}

\textbf{Auto-correlation.} All observed auto-correlations (aggregated over the 1000 repetitions) are negative or close to zero. There are some differences between variables, and the ACs of chain variances seem lower than the ACs of chain means. The general trend in figure \ref{fig:conv} is that ACs decrease steeply as the number of iterations increases.  \textbf{The lowest AC observed is %-0.735 
at four iterations. Simulation conditions with a higher number of iterations show a slow increase towards zero. Auto-correlations do not plateau completely within the range of simulation conditions considered in this study. It seems that ACs might become positive if more iterations would be allowed. The average absolute maximum ACs are negative up to 77 iterations for chain means, and for chain variances all iterations considered here have negative ACs.}
